Course Project 1 - Reproducible Research
--------------------------------------------------

## 1. Code for reading in the dataset and/or processing the data

```{r}
library(tidyverse)
activity_url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
temp <- tempfile()
download.file(activity_url, temp)

unzip(temp, "activity.csv")
mydata <- read.csv("activity.csv")
unlink(temp)
activity_omit <- na.omit(mydata)
total_ac_omit <- aggregate(steps ~ date, activity_omit, sum)

```


## 2. Histogram of the total number of steps taken each day

```{r}

ggplot(data = total_ac_omit,aes(x=steps)) +
        geom_histogram(bins = 20, fill = "darkblue", color = "black") +
        ggtitle("Total of steps per day") +
        ylab("Frequency") + xlab("Number of steps")

```

## 3. Mean and median number of steps taken each day


```{r}
steps_mean <- mean(total_ac_omit$steps)
steps_median <- median(total_ac_omit$steps)

print(steps_mean)

print(steps_median)

```

The mean is `steps_mean[1]` and the median is `steps_median[1]`


## 4. Time series plot of the average number of steps taken

Prepare data

```{r}
total_ac_average <- aggregate(steps ~ interval, activity_omit, mean)
```



```{r}

ggplot(total_ac_average, aes(interval, steps)) +
        geom_line() +
        ggtitle("Time series plot of the average number of steps taken") +
        xlab("Interval") +
        ylab("Number of steps")

```


## 5. The 5-minute interval that, on average, contains the maximum number of steps

```{r}
max_number_steps <- which.max(total_ac_average$steps)
print(total_ac_average[max_number_steps,])
```


## 6. Code to describe and show a strategy for imputing missing data

```{r}
fill <- function(steps, interval) {
        fill_ac <- NA
        if (!is.na(steps))
                fill_a <- c(steps)
        else
                fill_a <- (total_ac_average[total_ac_average$interval==interval, "steps"])
        return(fill_a)
}

new_activity <- mydata
new_activity$steps <- mapply(fill, new_activity$steps, new_activity$interval)

total_ac_steps <- aggregate(steps ~ date, new_activity, sum)

```


## 7. Histogram of the total number of steps taken each day after missing values are imputed

```{r}

ggplot(data = total_ac_steps,aes(x=steps)) +
        geom_histogram(bins = 20, fill = "darkblue", color = "black") +
        ggtitle("Total of steps per day") +
        ylab("Frequency") + xlab("Number of steps")


```

```{r}
total_ac_steps_mean <- mean(total_ac_steps$steps) 
total_ac_steps_median <- median(total_ac_steps$steps) 
print(total_ac_steps_mean)
print(total_ac_steps_median)
```


## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
week_activity <- new_activity

week_activity$date <- as.Date(strptime(week_activity$date, format="%Y-%m-%d")) 

week_activity$day <-  factor(ifelse(as.POSIXlt(week_activity$date)$wday %in% c(0,6), 'weekend', 'weekday'))
averaged_week_activity <- aggregate(steps ~ interval + day, data=week_activity, mean)

ggplot(averaged_week_activity, aes(interval, steps)) + 
        geom_line() + 
        facet_grid(day ~ .) +
        xlab("5-minute interval") + 
        ylab("avarage number of steps")

```


